---
title: "lab3"
author: "Merly Klaas & Cassie Malcom"
date: "1/31/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    highlight: kate
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE)
```

```{r}
library(pacman)

p_load(here, rio, tidyverse, colorspace, scico, ggplot2, gghighlight, colorblindr, viridis)

dat <- import(here("data","honeyproduction.csv"))
```

### 1. Visualize the total production of honey across years by state Use color to highlight the west coast (Washington, Oregon, and California) with a different color used for each west coast state.

  * Hint 1: I’m not asking for a specific kind of plot, just one that does the preceding. But if you’re trying to visualize change over time, a bar chart is likely not going to be the best choice.
  * Hint 2: To get each state to be a different color you should either map state to color (for your layer that adds the west coast colors) or use the gghighlight package.

```{r include=FALSE}
str(dat)
```

First Attempt - this plot with factors seems way too busy
```{r}
ggplot(dat, aes(state, prodvalue, color = state)) +
geom_col(alpha = 0.9) +
facet_wrap(~year)
```

2nd attempt - still difficult to discern the lines at the bottom
```{r}
ggplot(dat, mapping = aes(x = year, y = prodvalue, color = state)) +
    geom_line()
```

3rd attempt - Three states (WA, OR, CA) are highlighted in this version, but lose the color coding to discern all 44 states.
```{r}
ggplot(dat, mapping = aes(x = year, y = prodvalue)) +
    geom_line(aes(color = state), size = 1) + 
  gghighlight(state == "OR" | state == "WA"| state == "CA") +
  labs(title = "US Honey Production By State from 1998 to 2012", 
       x = "Years", y = "Honey Production") +  
  scale_y_continuous(expand = c(0,0), labels = scales::comma ) +
  theme_light()
```
  
### 2. Reproduce the plot according to three different kinds of color blindness, as well as a desaturated version.
```{r}
# devtools::install_github("wilkelab/cowplot")
# install.packages("colorspace", repos = "http://R-Forge.R-project.org")
# devtools::install_github("clauswilke/colorblindr")

# library(colorblindr)
```

```{r}
cbt <- ggplot(dat, mapping = aes(x = year, y = prodvalue)) +
    geom_line(aes(color = state), size = 1) + 
  gghighlight(state == "OR" | state == "WA"| state == "CA") +
  labs(title = "US Honey Production By State", 
       x = "Years", y = "Honey Production") 

colorblindr::cvd_grid(cbt)
```

### 3. Reproduce the plot using a color blind safe palette.

```{r}
cbud <- ggplot(dat, mapping = aes(x = year, y = prodvalue)) +
    geom_line(aes(color = state), size = 1) + 
  scale_color_viridis(discrete = TRUE) +
  gghighlight(state == "OR" | state == "WA"| state == "CA") +
  scale_y_continuous(labels = scales::comma ) +
    scale_x_continuous(expand = c(0,0)) +
  labs(title = "US Honey Production By State from 1998 to 2012", 
       x = "Years", y = "Honey Production") + 
  theme_light()
  

cbud
```

### 4. Download the file here denoting the region and division of each state.
  * Join the file with your honey file.
  * Produce a bar plot displaying the average honey for each state (collapsing across years).
  * Use color to highlight the region of the country the state is from.
  * Note patterns you notice.
  
```{r}
region <- import(here("data", "us census bureau regions and divisions.csv")) %>% 
  rename(state = `State Code`) 

new <- full_join(region, dat)

  new %>% 
    drop_na() %>% 
    group_by(Region, State) %>% 
    summarise(meanprod = mean(totalprod)) %>% 
    ggplot() +
    geom_col(aes(meanprod, fct_reorder(State, meanprod), fill = Region)) + 
      labs(x= "Average Honey Production",
       y = "State",
       title = "Avarage Honey Production by State from 1998 - 2012") + 
    theme_light()+
    scale_x_continuous(expand = c(0,0), labels = scales::comma ) + 
  theme (axis.ticks = element_blank()) + 
    scico::scale_fill_scico_d() 
   
```

Some emerging patterns:

  * Midwest Region relatively has highest honey production in contrast with Northeast Region that has relatively low honey production. 
  * There is no consistent pattern for South and West Regions as the variation of honey production is quite large per state within these regions.
  
### 5. Create a heatmap displaying the average honey production across years by region (averaging across states within region).
```{r}
new %>% 
   drop_na() %>% 
  group_by(year, Region) %>% 
  summarise(meanprod = mean(totalprod)) %>%
  ggplot()+ 
  geom_tile(aes(Region, year, fill = meanprod))+
  scale_fill_viridis_c(option = "magma", trans = "reverse") +
  coord_flip() +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        legend.key.size = unit(1, 'cm'),
        legend.key.height = unit(.5,"cm")) +
  labs(title = "Honey Production across year by Region",
       fill = "Production", 
       caption = "Notes: Data collected in 1998-2012; \nSource: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-05-21")


```

### 6. Create at least one more plot of your choosing using color to distinguish, represent data values, or highlight. If you are interested in producing maps, I suggest grabbing a simple features data frame of the US using the Albers projection by doing the following:

```{r}

map <- new %>% 
    group_by(State) %>% 
    summarise(meanprod = mean(totalprod)) 

remotes::install_github("hrbrmstr/albersusa")
library(albersusa)
us <- usa_sf()
join <- full_join(us, map, by=c("name" = "State"))

library(ggplot2)
ggplot(join) +
  geom_sf(aes(fill = meanprod)) +
  scale_fill_continuous_divergingx(palette = "Earth",  labels = scales::comma) +
  labs(title = "Honey Production across year by State from 1998 - 2012",
       fill = "Production", 
       caption = "Notes: Gray color is for states with missing values; Data collected in 1998-2012 \nSource: https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-05-21") +
  theme_void()
  
```




